---
title: "Data Analysis Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 



## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
 fatal_shootings<- read_csv("data/fatal-police-shootings-data.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```



### Question 1
*Are police shootings happening mostly in majority-black communities, even when the victim is not black?

```{r}

```

### Question 2
*Do these shootings mostly occur in low income areas? 

```{r}

```

### Question 3
*Are police shootings still happening disproportionately affecting black people regardless of the area (even in white neighborhoods)?

```{r}

```

### Question 4
*How many shootings occurred in neighborhoods where there is a school in the proximity?


```{r}

```

### Question 5
*Based off of #4, in the shootings that happened near the schools, were the victims more likely to be under 18?

```{r}


```

-30-