---
title: "Data Analysis Project"
author: "Kassidy, Maddie, Cassidie"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r}
install.packages("censusxy")
install.packages("tidycensus")
```

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
library(tidycensus)
```

```{r}
#census_api_key("cf1cf256ae2dbf70e349a8ed26a87792c4ad7275", install=TRUE)
```

## Load and Cleaning Data

```{r}
# Load required data
fatal_shootings<- read_csv("data/fatal-police-shootings-data.csv") 
# Clean required data and prepare for analysis if needed. 


```



### Question 1
*Are police shootings happening mostly in majority-black communities, even when the victim is not black?

```{r}
fatal_shootings <- fatal_shootings %>%
  fatal_shootings <- cxy_geocode(fatal_shootings, city = city, state = state, latitude = latitude, longitude = longitude, class = "sf", return = "geographies")
  


fatal_shootings<- fatal_shootings %>% 
    filter(!is.na(longitude)) 
shootings <- tibble()
for (row_number in 1:nrow(shootings)) {
  
  row_df<- shootings %>%
  slice(row_number)
   longitude <- row_df$longitude
  latitude <- row_df$latitude
  census_results <- cxy_geography(longitude, latitude) %>%
    select(Census.Tracts.GEOID) %>%
    clean_names()
 
   row_df <- row_df %>%
     bind_cols(census_results) %>%
  shootings <- shootings %>%
    bind_rows(row_df) 
   
   print(paste0("finished ", row_number, " ", Sys.time()))
   
   if (row_number%%500 == 0) {
     filepath <- paste0("data/geocoded_results_", row_number, ".rds")
     write_rds(shootings, filepath)
    shootings <- as_tibble()
   }
```

### Question 2
*Do these shootings mostly occur in low income areas? 

```{r}

```

### Question 3
*Are police shootings still happening disproportionately affecting black people regardless of the area (even in white neighborhoods)?

```{r}

```

### Question 4



```{r}

```

### Question 5


```{r}


```

-30-