---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**). 
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**). 
* write out the answer in the form of a complete sentence in the space given (look for **A**). 
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS. 

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse.To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}


# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data 

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data. 

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data. 
**Q1** How many rows and columns does this dataframe have?
**A1** This data frame has Rows: 33,703 Columns:63

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

**Q2** Which West Virginia county got the most loans? How many?
**A2** Kanawha County has the most loans of 4332.

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarize(
    count_loans =n()
  )%>%
  arrange(desc(count_loans))
```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** Wirt County has the fewest loans of 36.

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarize(
    count_loans =n()
  )%>%
  arrange(count_loans)

```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** It is probably because the wealth gap in these two counties. A county that is richer does  not need more laons.


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** The Greenbrier Hotel Corporation got the highest loan of 8912202. It is a hotel business.It retained 500 jobs. 

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name,name,business_type,amount,jobs_retained) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount)
  ) %>%
  arrange(desc(total_loans_amount))
   
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 2? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** Govenor Jim Justice who is West Virginia's richest man.

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings. 

**A7** I found that statewide the total loan amount was 2606525203. This is a huge number and I'd be interseted in comparing it to other states to see if its more or less. I don't think West Virginia is a hugely wealthy state so I think that would make an interesting point in a story. 
```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(state) %>%
  summarize(
   count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount)
  ) %>%
  arrange(desc(mean_loan_amount))
```

**Q8** Produce a summary table that has the following information for each West Virginia county:
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_county_summary` so that you can examine it with the data viewer. 

What newsworthy nuggets or patterns jump out at you that you could imagine dropping into a news story or doing follow up reporting on? Write up a few sentence that summarizes one county. 

**A8** The county I chose is Calhoun which had the highest mean loan account. I thought that was intersting because for the rest of the categories it was in the middle of all the other counties- not high or low for the rest of the categories of count, total, median. I would want to look into why it has the higest mean loan and what exactly does that say about the county in general. 
```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarize(
   count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount)
  ) %>%
  arrange(desc(mean_loan_amount))
```
